---
title: "scATAC-seq Rat Metrial Glands"
author: "Ha T. H. Vu"
output: html_document
---
  
```{r setup, include=FALSE}
options(max.print = "75")
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "Files/",
  fig.width = 15,
  prompt = FALSE,
  tidy = FALSE,
  message = FALSE,
  warning = TRUE
)
knitr::opts_knit$set(width = 75)
```

This is a documentation for analyses of scATAC-seq data, generated from rat metrial gland tissues on gestational day (GD) 15.5 and 19.5. <br>
  
## Conserved common peak - conserved common gene networks:
Build list of genes that are TBC markers at both gd15.5 and gd19.5 that are conserved in human EVT using scRNA-seq data, and these markers are also near to the common peaks.
```{r}
marker15.5 <- read.table("/work/LAS/geetu-lab/hhvu/project3_scATAC/scATAC-seq-analysis/4_RNAintegration/scRNA-seqFiles/gd15.5-wilcoxin-cluster23.txt", header = T, sep = "\t")
marker19.5 <- read.table("/work/LAS/geetu-lab/hhvu/project3_scATAC/scATAC-seq-analysis/4_RNAintegration/scRNA-seqFiles/gd19.5-wilcoxin-cluster6.txt", header = T, sep = "\t")
markers <- intersect(intersect(marker15.5$gene, marker19.5$gene), closest_genes$gene_name)

conservedRats <- read.table("/work/LAS/geetu-lab/hhvu/project3_scATAC/scATAC-seq-analysis/5_integrating2timepoints/annotation_scRNAseq/commonPeaks/conservedRat.txt", header = T, sep = "\t")

markers <- markers[which(toupper(markers) %in% conservedRats$ratGenes == T)]

length(markers)
```

Convert common peaks (gd19.5 coordinates) to human coordinates using LiftOver, then find the regions that have ATAC-seq peaks in EVT. EVT peaks were obtained from https://www.medrxiv.org/content/10.1101/2022.05.25.22275520v1.full-text
```{Bash, eval = F}
cd /work/LAS/geetu-lab/hhvu/project3_scATAC/scATAC-seq-analysis/5_integrating2timepoints/annotation_scRNAseq/commonPeaks
module load bedtools2
sed -i 's/chr//g' commonPeaks-gd19.5-hg38Coor.bed
bedtools intersect -a commonPeaks-gd19.5-hg38Coor.bed -b EVTpeaks.bed -wa | sort -u > commonPeaks-gd19.5-hg38Coor-inEVT.bed
```

Find peaks with enriched motifs and are also conserved
```{r}
motifPeaks <- read.table("/work/LAS/geetu-lab/hhvu/project3_scATAC/scATAC-seq-analysis/5_integrating2timepoints/annotation_scRNAseq/commonPeaks/commonPeaksGD19.5withFinalMotifs.txt", header = T)
conservedPeaks <- read.table("/work/LAS/geetu-lab/hhvu/project3_scATAC/scATAC-seq-analysis/5_integrating2timepoints/annotation_scRNAseq/commonPeaks/commonPeaks-gd19.5-hg38Coor-inEVT.bed", header = F)
motifPeaks <- motifPeaks[motifPeaks$peak %in% conservedPeaks$V4,]
#write.table(motifPeaks, "/work/LAS/geetu-lab/hhvu/project3_scATAC/scATAC-seq-analysis/5_integrating2timepoints/annotation_scRNAseq/commonPeaks/conservedCommonPeaksGD19.5withFinalMotifs.txt", sep = "\t", row.names = F, quote = F)
```

Number of pekas with enriched motifs and are also conserved
```{r}
length(unique(motifPeaks$peak))
```

Associate peaks/motifs to genes
```{r}
network <- inner_join(closest_genes, motifPeaks, by = c("query_region" = "peak"))
network <- inner_join(network, motifs19.5[,c("motif", "motif.name")], by = c("motif" = "motif"))
network <- network[network$gene_name %in% markers,]
dim(network)

network <- distinct(network)

write.table(network, "/work/LAS/geetu-lab/hhvu/project3_scATAC/scATAC-seq-analysis/5_integrating2timepoints/annotation_scRNAseq/commonPeaks/conservedCommonPeaksGenes-gd19.5-motifLoc.txt", sep = "\t", row.names = F, quote = F)
```

Grouping motifs of the same gene families together. Here if two families shared an element, I will also merge them together.
```{r}
jaspar <- read.table("/work/LAS/geetu-lab/hhvu/JASPAR-library.txt", header = T, sep = "\t")
jaspar <- jaspar[jaspar$ID %in% network$motif,]
network <- inner_join(network, jaspar[,c("ID", "Family")], by = c("motif" = "ID"))
keep <- network$Family
keep <- subColon(keep)
for (i in 1:length(keep)) {
  keep[i] <- list(unique(keep[i][[1]]))
  for (j in setdiff(1:length(keep), which(keep %in% keep[i]))) {
    if (sum(keep[j][[1]] %in% keep[i][[1]]) > 0) {
      keep[j] <- keep[i]
    }
  }
}
keep2 <- unlist(lapply(keep, toString))
network$Family2 <- keep2

all <- data.frame(tfGene=character(), regGene=character())
for (i in unique(network$Family2)) {
  tfGene <- unique(network[which(network$Family2 == i), "motif.name"])
  tf <- toString(tfGene)
  regGene <- unique(network[which(network$Family2 == i), "gene_name"])
  tbl <- data.frame(matrix(nrow = length(tfGene)*length(regGene), ncol = 2))
  for (j in 1:length(tfGene)) {
    tbl[((j-1)*length(regGene)+1):(j*length(regGene)),1] <- tf
    tbl[((j-1)*length(regGene)+1):(j*length(regGene)),2] <- as.character(regGene)
  }
  all <- rbind(all, tbl)
}
all$X2 <- toupper(all$X2)
all <- dplyr::distinct(all)

g <- unique(all$X2)
for (k in 1:length(g)) {
  if (length(grep(g[k], all$X1)) != 0) {
    all[which(g[k] == all$X2), "X2"] <- all[grep(g[k], all$X1)[1], "X1"]
  }
}
all <- dplyr::distinct(all)

#write.table(all, "/work/LAS/geetu-lab/hhvu/project3_scATAC/scATAC-seq-analysis/5_integrating2timepoints/annotation_scRNAseq/commonPeaks/conservedCommonPeaksGenes-gd19.5.txt", sep = "\t", row.names = F, quote = F)
```
