---
title: "scATAC-seq Rat Metrial Glands"
author: "Ha T. H. Vu"
output: html_document
---
  
```{r setup, include=FALSE}
options(max.print = "75")
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "Files/",
  fig.width = 15,
  prompt = FALSE,
  tidy = FALSE,
  message = FALSE,
  warning = TRUE
)
knitr::opts_knit$set(width = 75)
```

This is a documentation for analyses of scATAC-seq data, generated from rat metrial gland tissues on gestational day (GD) 15.5 and 19.5. <br>
In order to carry out motif analysis, we need a `BSgenome` library for Rnor6. Following the instructions from https://www.bioconductor.org/packages/release/bioc/vignettes/BSgenome/inst/doc/BSgenomeForge.pdf we can develop the library. <br>

First, we need to obtain DNA sequence fasta files from Ensembl. The fasta files of each chromosome should be named as `chromosomeName.fa`. For example, `1.fa` for DNA sequence of chromosome 1. <br>
Second, create a seed file like the following:
```
Package: BSgenome.Rnorvegicus.Ensembl.rn6
Title: Full genome sequences for Rattus norvegicus (Ensembl archive Sept. 2020)
Description: Full genome sequences for Rattus norvegicus as provided by Ensembl, Sept. 2020 version.
Version: 0.0.1
organism: Rattus norvegicus
common_name: Rat
provider: Ensembl
provider_version: rn6
release_date: Sept. 2020
release_name: rnor6
source_url: ftp://ftp.ensembl.org/pub/release-98/fasta/rattus_norvegicus/
organism_biocview: Rattus_norvegicus
BSgenomeObjname: Rnorvegicus
seqnames: c(1:20, "X", "Y")
seqs_srcdir: /work/LAS/geetu-lab/hhvu/project3_scATAC/rnor6-Ensembl
```
Third, build the package in R:
```{r, eval=FALSE}
library(BSgenome)
forgeBSgenomeDataPkg("/work/LAS/geetu-lab/hhvu/project3_scATAC/rnor6-Ensembl/BSgenome.Rnorvegicus.Ensembl.rn6-seed")
```
Last, in command line, do the following:
```{BASH, eval=FALSE}
R CMD build /work/LAS/geetu-lab/hhvu/project3_scATAC/rnor6-Ensembl/
R CMD check /work/LAS/geetu-lab/hhvu/project3_scATAC/rnor6-Ensembl/
R CMD INSTALL /work/LAS/geetu-lab/hhvu/project3_scATAC/rnor6-Ensembl/
```
After this, the package `BSgenome.Rnorvegicus.Ensembl.rn6` will be ready to use.

```{r}
ssessionInfo()
```